## Functions and Conditionals

### Conditionals

The basic syntax for an if statement goes like this:

```{r}
if (logical statement) {
  do something
} else {
  do something else
}
```

The logical statements should boil down to a boolean; for example:

```{r}

```

We can also add else if statements in the middle if we need more options.

```{r}

```

Neat! We can also nest conditionals just like we do in Python.

```{r}

    
```

Neat! Now, let's look at another core part of R programming - Loops!

### Loops

```{r}

```

We can also do nifty stuff with our vectors, such as apply changes to specific indices of our data.

```{r}

```

We can loop over more than just lists:

```{r}

```

Like Python, we have while loops in R:

```{r}

```

## Functions

When should we use functions?

```{r}
our_function <- function(optional_params){
  # do some stuff
}
```

An example:

```{r}

```

This is all cool.. but it didn't do anything. We need to call the function!

```{r}

```

Notice that in the background here, we are looping over each item in the array, but we never write a for loop! Keep this in mind, since when you pass in a large data set, the changes made by functions aren't quite immediate - the computer does have to go over each item in your data set, and change them one at a time.

A shortcut of sorts is to use the `apply()` function, which operates very similarly to Python's `map()`. Note that `apply` syntax goes as follows: `apply(X, MARGIN, FUN)` where X is the data set to use, MARGIN is specification for rows or columns, and FUN is the function to apply.

```{r}

```

It is common to use `mean`, `median`, `sum`, `min`, or `max` with `apply`, but we can also use our user-defined functions as well. Note that apply() only works with matrices.

```{r}

```

### Homework

1.  Do 5 CodeWars problems in R. Share your solutions here:

    ```{r}
    #1 Complete the solution so that it reverses all of the words within the string passed in.Words are separated by exactly one space and there are no leading or trailing spaces.

    solution <- function(sentence){
      string_split = strsplit(as.character(sentence), split = " ")
      string_length = length(string_split[[1]])
      if (string_length == 1) {
        reversed_string = string_split[[1]]
      } else {
        reversed_split = string_split[[1]][string_length:1]
        reversed_string = paste(reversed_split, collapse = " ")
      }
      return(reversed_string)
    }

    #2 Given two integers a and b, which can be positive or negative, find the sum of all the integers between and including them and return it. If the two numbers are equal return a or b.

    get_sum <- function(a, b){
      if (a == b) {
        result <- a
      } else {
        if (a<b) {
          result <- sum(a:b)
        } else {
          result <- sum(b:a)
        }
      }
      return(result)
    }

    #3 Finish the solution so that it sorts the passed in array of numbers. If the function passes in an empty array or null/nil value then it should return an empty array.

    solution <- function(nums) {
      sort(nums)
    }

    #4 Nathan loves cycling.Because Nathan knows it is important to stay hydrated, he drinks 0.5 litres of water per hour of cycling.You get given the time in hours and you need to return the number of litres Nathan will drink, rounded to the smallest value.

    litres <- function(time) {
      floor(time*.5)
    }

    #5 Write a function that takes a list of strings as an argument and returns a filtered list containing the same elements but with the 'geese' removed.

    geese <- c("African", "Roman Tufted", "Toulouse", "Pilgrim", "Steinbacher")

    goose_filter <- function(birds){
            birds[!birds%in%geese]
    }

    ```

2.  Create a function that checks an input of type `int` and returns fizz if the number is a multiple of 3, buzz if the number is a multiple of 5, and fizzbuzz if the number is a multiple of 3 and 5.

    ```{r}
    for(i in 1:100){
      if(i%%3==0 & i%%5==0){
        print('FizzBuzz')
      }
      else if(i%%3==0){
        print('Fizz')
      }
      else if(i%%5==0){
        print('Buzz')
      }
      else{
        print(i)
      }
    }

    ```

3.  **Mario Challenge:** In the classic Super Mario games, Mario ends each level by ascending a flight of stairs and leaping off the top to arrive at the castle at the end. (See [here](https://youtu.be/-avspZlbOWU?t=55) for details). We can imagine that each of the bricks of the stairs is a character X. Create a function that prints out a set of stairs. For example:

    If the function is called marioStairs() and we pass in the value 4, we would expect to see stairs with a length of 4 and a height of 4:

    X

    XX

    XXX

    XXXX

    ```{r}
    # You may feel more comfortable figuring this out in Python first, then translating into R later.

    marioStairs <- function(){
      for(i in 1:4){
        print(noquote(replicate(i,'X')))
      }
    }

    marioStairs()
    ```
